---
title: "DSC 520 Final Project"
author: "Bilal Kudaimi"
date: "2020-11-07"
output: 
  pdf_document: default
---

# Summary of the data

```{r setup, echo = FALSE}
library('ggplot2')
library('tidyr')
library('dplyr')
library('car')
library('tm')
options(warn=-1)

#Importing the data and checking for NA values
cars <- read.csv("C:/Users/PS3ma/Downloads/ThinkStats2-master/ThinkStats2-master/code/CO2 Emissions_Canada.csv")
cars$Transmission <- removeNumbers(cars$Transmission)
empty_data <- is.na(cars)

#Introducing noise to account for rounding
cars$Engine.Size.L. <- jitter(cars$Engine.Size.L., amount = 0.25)
cars$CO2.Emissions.g.km. <- jitter(cars$CO2.Emissions.g.km., amount = 0.5)

#Transforming all categorical variables into factors
cars$Transmission <- as.factor(cars$Transmission)
cars$Fuel.Type <- as.factor(cars$Fuel.Type)
cars$Make <- as.factor(cars$Make)
cars$Vehicle.Class <- as.factor(cars$Vehicle.Class)

#Summary of the dataset
summary(cars)
str(cars)


```

# Histograms and bar charts of engine specifications to find the most common engine type

```{r histograms, echo = FALSE}
#Histograms and bar charts of engine specifications to find the most common engine type
ggplot(cars, aes(x = Engine.Size.L.)) + geom_histogram(bins = 25) + xlab('Engine Displacement (L)') + ylab('Frequency')
ggplot(cars, aes(x = Cylinders)) + geom_histogram() + xlab('Number of Cylinders') + ylab('Frequency')
ggplot(cars, aes(x = Transmission)) + geom_bar() + ylab('Transmission Type') + ylab('Frequency')
ggplot(cars, aes(x = Fuel.Type)) + geom_bar() + xlab('Fuel Type') + ylab('Frequency')

```

# Scatter plots and box plots to find what the correlation is between each variable and CO2 emissions

```{r scatter-and-box, echo = FALSE}
ggplot(cars, aes(x = reorder(Make, CO2.Emissions.g.km.), y = CO2.Emissions.g.km.)) + geom_boxplot(outlier.shape = NA) + theme(axis.text.x=element_text(angle = 90)) + xlab('Make') + ylab('CO2 Emissions (g/km)')
ggplot(cars, aes(x = reorder(Vehicle.Class, CO2.Emissions.g.km.), y = CO2.Emissions.g.km.)) + geom_boxplot(outlier.shape = NA) + theme(axis.text.x=element_text(angle = 90)) + xlab('Vehicle Shape') + ylab('CO2 Emissions (g/km)')
ggplot(cars, aes(x = Engine.Size.L., y = CO2.Emissions.g.km.)) + geom_point(color = "blue") + xlab('Engine Displacement (L)') + ylab('CO2 Emissions (g/km)')
ggplot(cars, aes(x = reorder(as.factor(Cylinders), CO2.Emissions.g.km.), y = CO2.Emissions.g.km.)) + geom_boxplot(outlier.shape = NA) + xlab('Cylinders') + ylab('CO2 Emissions (g/km)')
ggplot(cars, aes(x = reorder(Transmission, CO2.Emissions.g.km.), y = CO2.Emissions.g.km.)) + geom_boxplot(outlier.shape = NA) + xlab('Transmission Type') + ylab('CO2 Emissions (g/km)')
ggplot(cars, aes(x = reorder(Fuel.Type, CO2.Emissions.g.km.), y = CO2.Emissions.g.km.)) + geom_boxplot(outlier.shape = NA) + xlab("Fuel Type") + ylab('CO2 Emissions (g/km)')
ggplot(cars, aes(x = Fuel.Consumption.City..L.100.km., y = CO2.Emissions.g.km.)) + geom_point(color = "green") + xlab('City Fuel Consumption (L/100km)') + ylab('CO2 Emissions (g/km)')
ggplot(cars, aes(x = Fuel.Consumption.Hwy..L.100.km., y = CO2.Emissions.g.km.)) + geom_point(color = "red") + xlab('Hwy Fuel Consumption (L/100km)') + ylab('CO2 Emissions (g/km)')
ggplot(cars, aes(x = Fuel.Consumption.Comb..L.100.km., y = CO2.Emissions.g.km.)) + geom_point(color = "purple") + xlab('Comb Fuel Consumption (L/100km)') + ylab('CO2 Emissions (g/km)')

```


## The emissions vs fuel consumption plot displayed multiple lines. Let's separate that data to see if we can see any interesting information.

## Starting with separating by fuel type:


```{r fuel-color, echo = FALSE}

#Separating the data frame by transmission type 
cars_A <- filter(cars, Transmission == 'A')
cars_AM <- filter(cars, Transmission == 'AM')
cars_AS <- filter(cars, Transmission == 'AS')
cars_AV <- filter(cars, Transmission == 'AV')
cars_M <- filter(cars, Transmission == 'M')

#Color coding the emissions vs fuel consumption plot by fuel type. We can clearly see
#that each of the lines seen in the original plot is due to a different fuel type.
ggplot(cars, aes(x = Fuel.Consumption.Comb..L.100.km., y = CO2.Emissions.g.km., color = Fuel.Type)) + geom_point() + xlab('Comb Fuel Consumption (L/100km)') + ylab('CO2 Emissions (g/km)')

```

## We can clearly see that each of the lines seen in the original plot is due to a different fuel type. 

## Now I will separate by transmissions and keep the fuel type color coding.

```{r transmission-color, echo = FALSE}

#Separating by transmission and color coding by fuel type reveals that some transmissions appear to be more common with some fuels.
ggplot(cars_A, aes(x = Fuel.Consumption.Comb..L.100.km., y = CO2.Emissions.g.km., color = Fuel.Type)) + geom_point() + xlab('Comb Fuel Consumption (L/100km)') + ylab('CO2 Emissions (g/km)')
ggplot(cars_AM, aes(x = Fuel.Consumption.Comb..L.100.km., y = CO2.Emissions.g.km., color = Fuel.Type)) + geom_point() + xlab('Comb Fuel Consumption (L/100km)') + ylab('CO2 Emissions (g/km)')
ggplot(cars_AS, aes(x = Fuel.Consumption.Comb..L.100.km., y = CO2.Emissions.g.km., color = Fuel.Type)) + geom_point() + xlab('Comb Fuel Consumption (L/100km)') + ylab('CO2 Emissions (g/km)')
ggplot(cars_AV, aes(x = Fuel.Consumption.Comb..L.100.km., y = CO2.Emissions.g.km., color = Fuel.Type)) + geom_point() + xlab('Comb Fuel Consumption (L/100km)') + ylab('CO2 Emissions (g/km)')
ggplot(cars_M, aes(x = Fuel.Consumption.Comb..L.100.km., y = CO2.Emissions.g.km., color = Fuel.Type)) + geom_point() + xlab('Comb Fuel Consumption (L/100km)') + ylab('CO2 Emissions (g/km)')
#Automatic and hybrid transmissions can be found on all car types, manual and automated manual transmissions are more common in premium gas cars, and CVTs are more common in regular gas cars. This may be because sports cars, which use premium gas, tend to use manually controlled transmissions due to their powerful engines, whereas CVTs are meant to reduce pollution and fuel consumption so cars that have them use regular gas, which emits the least CO2 of the fuels.
```

## This separation reveals that some transmissions appear to be more common with some fuels.

## Automatic and hybrid transmissions can be found on all car types, manual and automated manual transmissions are more common in premium gas cars, and CVTs are more common in regular gas cars. This may be because sports cars, which use premium gas, tend to use manually controlled transmissions due to their powerful engines, whereas CVTs are meant to reduce pollution and fuel consumption so cars that have them use regular gas, which emits the least CO2 of the fuels.


# Now I conduct two sample t tests on the following:

> The combined fuel consumptions of gas and premium gas cars to see if there is a difference

> The CO2 emissions of gas and diesel cars to determine if diesel pollutes more

> The CO2 emissions of gas ethanol cars to determine if ethanol is more eco-friendly

## Based on the t tests, there is a significant difference between gas and premium gas fuel consumption. There is also a significant difference between gas and diesel CO2 emissions, and gas and ethanol CO2 emissions. More data on diesel and ethanol cars will need to be obtained in the future, since their sample sizes are very small.

``` {r t-tests, echo = FALSE}
cars_gas <- filter(cars, Fuel.Type == 'X')
cars_premium_gas <- filter(cars, Fuel.Type == 'Z')
cars_diesel <- filter(cars, Fuel.Type == 'D')
cars_ethanol <- filter(cars, Fuel.Type == 'E')


#A small difference was seen between the gas and premium gas lines in the scatter plot. 
#Is there a difference between gas and premium gas fuel consumption? Let's find out.
res <- t.test(cars_gas$Fuel.Consumption.Comb..L.100.km., cars_premium_gas$Fuel.Consumption.Comb..L.100.km.)
res
#Based on the p-value, there is indeed a significant different in fuel consumption between gas and premium gas cars.


#Do diesel cars really pollute more than regular gas cars? Let's find out as well.
res2 <- t.test(cars_gas$CO2.Emissions.g.km., cars_diesel$CO2.Emissions.g.km.)
res2
#Based on the p-value of 0.46, the difference in emissions between gas and diesel is barely significant. This appears to confirm the common claim that diesel cars pollute more despite being more fuel efficient; however, the diesel sample is an order of magnitude smaller than the gas sample so more data will be needed to confirm this.

#Ethanol cars appear to have a higher mean emissions rate than other fuels, despite people claiming 
#ethanol to be a clean burning fuel. Is there an emission difference between ethanol and normal gas?
res3 <- t.test(cars_gas$CO2.Emissions.g.km., cars_ethanol$CO2.Emissions.g.km.)
res3
#It looks like ethanol does appear to have a significant difference from gas in emissions. The box plot reveals that ethanol has higher emissions than gas. This could be caused by the number of ethanol samples being an order of magnitude smaller than the number of gas samples. More ethanol samples will be needed to confirm these results.


```



# The scatter plots and box plots revealed that all the variables are linearly correlated with CO2 emissions. Therefore, I will set up a multiple linear regression model to predict CO2 emissions.

``` {r model, echo = FALSE}
#It appears that all the variables are linearly correlated with CO2 emissions so I
#will set up a multiple linear regression model on the data.
emission.model <- lm(CO2.Emissions.g.km. ~ Make + Vehicle.Class + Transmission + Fuel.Type + Cylinders + Engine.Size.L. + Fuel.Consumption.City..L.100.km. + Fuel.Consumption.Hwy..L.100.km. + Fuel.Consumption.Comb..L.100.km., data = cars)

summary(emission.model)

#Calculating R2
summary(emission.model)$r.squared


```

## Calculating the RMSE of the model

``` {r rmse, echo = FALSE}

#Calculating the RMSE
cars$standardized.residuals <- rstandard(emission.model)
RMSE <- sqrt(var(cars$standardized.residuals))
RMSE
```

## All Cook's distances are less than 1, so nothing is having an undue influence on the model.
## The following formula was used to find the bounds the covariance ratio should be in:

$1\pm[3(k+1)/n]$

## About 5% of ratios are outside the boundaries (shown below), but because all Cook's distances are below 1, there is little cause for alarm.
 

``` {r diagnostics, echo = FALSE}
#Casewise diagnostics (including standardized residuals, leverage, cook distance, and covariance ratios)
cars$standardized.residuals <- rstandard(emission.model)
cars$cooks.distance <- cooks.distance(emission.model)
cars$leverage <- hatvalues(emission.model)
cars$covariance.ratios <- covratio(emission.model)

#All Cook's distances are less than 1, so nothing is having an undue influence on the model
cars$cooks.distance[cars$cooks.distance > 1]


#The formula 1+/-[3(k+1)/n] was used to find the bounds the covariance ratio should be in 
cars$covariance.ratios[cars$covariance.ratios > 1.031 | cars$covariance.ratios < 0.969]
#About 5% are outside the boundaries, and because the Cook's distance is below 1, there is little cause for alarm.

```


## To test the assumption of independece, I conduct a Durbin-Watson test on the model. Because the D-W statistic is above 1, the assumption has been met, indicating the residuals are correlated.

``` {r, echo = FALSE}
#The assumption of independence has been met, as the D-W statistic is above 1, indicating the residuals are correlated.
durbinWatsonTest(emission.model)
```

## To determine if the model contains multicollinearity, I will construct a correlation matrix and calculate the determinant. If the determinant is greater than 0.00001, then multicollinearity is not a problem. 

## The determinant was found to be significantly smaller than this. This makes sense as many variables were highly linearly correlated with CO2 emissions, and thus, each other.


``` {r, echo = FALSE}
#To determine if the model contains multicollinearity, I will construct a correlation matrix and calculate the determinant. If the determinant is greater than 0.00001, then multicollinearity is not a problem. However, the determinant was found to be significantly smaller than this. This makes sense as many variables were highly correlated with CO2 emissions, and thus, each other. To find the determinant, the categorical variables are removed, the cor() function is then applied to calculate the correlation matrix, then the determinant function is then applied.
cor_matrix <- cor(select(cars, -Transmission, -Fuel.Type, -Make, -Model, -Vehicle.Class))
det(cor_matrix)
```

## The histogram and Q-Q plot inform us of any deviations from normality seen in the residuals. Based on this, we can clearly see that the distribution of residuals is approximately normal. In addition, the plot of residuals vs fitted values appears to be not random. Residuals appear to remain fairly constant across the fitted values. Due to the non-random residual vs fitted value distribution, the normality of residuals, and the assumption of independence, this model has little bias.

```{r, echo = FALSE}
#The histogram and Q-Q plot inform us of any deviations from normality seen in the residuals. Based on this, we can clearly see that the distribution of residuals is approximately normal. In addition, the plot of residuals vs fitted values appears to be not random. Residuals appear to remain fairly constant across the fitted values. Due to the non-random residual vs fitted value distribution, the normality of residuals, and the assumption of independence, this model has little bias.

hist(cars$standardized.residuals)

plot(emission.model)


```

## CO2 emissions were highly linearly correlated with fuel consumption, so it is safe to assume that the same variables most influential on emissions will influence fuel consumption the most as well.
