---
title: "Housing"
author: "Brett Sutow"
date: "10/14/2020"
output:
  word_document: default
  pdf_document: default
---
## Set the working directory to the root of your DSC 520 directory
##Set dataframe
```{r setup, include=FALSE}
library(readxl)
setwd("/Users/Brett/Desktop/dsc520-master")

housing_df<- read_excel('/Users/Brett/Downloads/week-6-housing.xlsx')
head(housing_df)
tail(housing_df)
summary(housing_df)
saleprice<- housing_df[,c(2)]
sqfoot<- housing_df[,c(22)]
bedroom<- housing_df[,c(15)]
```
1. Cleaning Datapoints based off the information and the given datapoints we would clean: City Name, Property Type. These have information that is not full, and not relevant. 


##Creating Variables: Sale date picked as another predictor because it allows for comparison based on time

```{r}
sale_lm <- lm(housing_df$`Sale Price` ~ housing_df$sq_ft_lot)
summary(sale_lm)
saleonbed_lm <- lm(housing_df$`Sale Price` ~ housing_df$sq_ft_lot + housing_df$bedrooms)
summary(saleonbed_lm)

```

## Including Plots

2. Rsquared represents the correlation between the dependent variable and the independent variable. In this case it shows for the sale_lm an rsquared of .01435 and the rsquared for the model saleonbed_lm .06278. The addition of the sale year helped improve the rsquared.Showing a better correlation based off that, but rsquare is still very low for us to determine impact between the variables. 
##Beta
```{r}
library(lm.beta)
lm.beta(sale_lm)
lm.beta(saleonbed_lm)
```

3. #The beta coefficients in for these datapoints shows a positive relationship.This tells us that as the independent variable increases so does the dependent.These betas are not very strong, as shown in the results. But, they are positive in result.


##Confidence Interval
```{r}
confint(sale_lm)
confint(saleonbed_lm)
```

4. We are 97.5% confident that our data points from the created variables will fall within the ranges listed in the charts produced.

##ANova
```{r}
anova(sale_lm)
anova(saleonbed_lm)
```
5.The residuals decreased from the original variable to the new created one.This means that means there is less of a relationship due to this addition.

##Residuals Output

```{r}
function(sale_lm)
function(saleonbed_lm)
residsale_lm<- resid(sale_lm)
residsaleonbed_lm<-resid(saleonbed_lm)
rstansale_lm<-rstandard(sale_lm)
rstansaleonyear_lm<-rstandard(saleonbed_lm)
print(rstansale_lm)
print(residsaleonbed_lm)
sum(residsale_lm, na.rm = TRUE)
sum(residsaleonbed_lm, na.rm= TRUE)
```

##Leverage,cooks distance, covariance

```{r}
hatvalues(sale_lm)
hatvalues(saleonbed_lm)
cooks.distance(sale_lm)
cooks.distance(saleonbed_lm)
cov(housing_df$`Sale Price`, housing_df$sq_ft_lot)
cov(housing_df$`Sale Price`,housing_df$bedrooms)
```
##Indpendence and Multiccollinearity

```{r}
library(MASS)
tbl= table(housing_df$`Sale Price`, housing_df$sq_ft_lot)
tbl1= table(housing_df$`Sale Price`, housing_df$bedrooms)
chisq.test(tbl)
chisq.test(tbl1)
vif(saleonbed_lm)

```

##Plot
```{r}
plot(residsale_lm)

plot(residsaleonbed_lm)

hist(residsale_lm)
hist(residsaleonbed_lm)
```
6. Based off the information above I believe the data is not biased. Though one of the key factors we can look at is  multicollinearity and it is above 1.It is not above it substainally enough to show significant impact on the data. 