---
title: "Exercise14SutowBrett"
author: "Brett Sutow"
date: "10/24/2020"
output: word_document
---
1. Based off what we see in the accuracy test in the target_category and test_category it shows that the accuracy is off. We can see these when we compare the various numbers in the charts. We are getting higher numbers in the test

2. We can see in the output comparison between accuracy and KNN, that there are still differences.Meaning this shows that the data does not match fully with the predictive model. 

3. The main difference is that KNN will always stay the same in output, where as from the logistic regression we can see changes throughout because we can see it is using random values. .3333 Would have little value for accuracy. 


```{r setup, include=FALSE}
setwd("/Users/Brett/Desktop/dsc520-master")
binary_df<- read.csv("/Users/Brett/Desktop/dsc520-master/data/binary-classifier-data.csv")
head(binary_df)
binarylrm<- glm(binary_df$label ~ binary_df$x + binary_df$y)
binarylrm
summary(binarylrm)

```



```{r, echo=FALSE}
normalize <-function(x) { (x -min(x))/(max(x)-min(x))   }
binary_norm <- as.data.frame(lapply(binary_df[,c(2:3)],normalize))
head(binary_norm)

data_split <- sample(1:nrow(binary_norm), .8 * nrow(binary_norm))
train<- binary_norm[data_split,]
head(train)
dim(train)

test<- binary_norm[-data_split,]

target_category<-binary_df[data_split, 3]
test_category<- binary_df[-data_split, 3]

target_category
test_category
```


```{r KNN, echo=FALSE}
library(FNN)
test_pred<- knn(train, test, cl=target_category)
test_pred

df_pred= data.frame(test_category,test_pred)
df_pred

table<- table(test_category, test_pred)
table
```


```{r, echo=FALSE}
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}
accuracy(table)
```


