---
title: "Exercise15"
author: "Brett Sutow"
date: "10/30/2020"
output: word_document
---
#Setup
```{r, echo= FALSE}
setwd("/Users/Brett/Desktop/dsc520-master")
binary_df<- read.csv("/Users/Brett/Desktop/dsc520-master/data/binary-classifier-data.csv")
trinary_df<- read.csv("/Users/Brett/Desktop/dsc520-master/data/trinary-classifier-data.csv")
library(FNN)
library(ggplot2)
```

## Answer 1
a.Plots below:


```{r echo= FALSE}
plot(binary_df$x, binary_df$y)
ggplot( binary_df, aes(x= x, y= y)) + geom_point( )+
  labs(title= "X vs Y", x="X", y="Y")

plot(trinary_df$x, trinary_df$y)
ggplot( trinary_df, aes(x= x, y= y)) + geom_point( )+
  labs(title= "X vs Y", x="X", y="Y")
```

## Answer 2

b. Below are the answers for KNN with accuracy for each value of the k. It also shows graphs of the plot where x-axis is the different values of k and y-axis is the accuracy model. Please see below results:

## Answer 3
c. I believe that linear classifiers would work for both the binary and trinary plots. Looking at the data plotted it shows that there would be success using the linear classifiers. Through my understanding these plots would work with a linear type model.

## Answer 2 Results:
```{r pressure, echo= FALSE}
normalize <-function(x) { (x -min(x))/(max(x)-min(x))   }
binary_norm <- as.data.frame(lapply(binary_df[,c(2:3)],normalize))
head(binary_norm)

data_split <- sample(1:nrow(binary_norm), .8 * nrow(binary_norm))
train<- binary_norm[data_split,]
head(train)
dim(train)

test<- binary_norm[-data_split,]

target_category<-binary_df[data_split, 2]
test_category<- binary_df[-data_split, 2]

target_category
test_category

binary3<- knn(train, test,cl=target_category, k=3)
binary3

binary5<- knn(train, test,cl=target_category, k=5)
binary5

binary10<- knn(train, test,cl=target_category, k=10)
binary10

binary15<- knn(train, test,cl=target_category, k=15)
binary15

binary20<- knn(train, test,cl=target_category, k=20)
binary20

binary25<- knn(train, test,cl=target_category, k=25)
binary25

tri_norm <- as.data.frame(lapply(trinary_df[,c(2:3)],normalize))
head(tri_norm)

data_split <- sample(1:nrow(tri_norm), .8 * nrow(tri_norm))
train2<- tri_norm[data_split,]
head(train2)
dim(train2)

test2<- tri_norm[-data_split,]

target_category1<-trinary_df[data_split, 3]
test_category1<- trinary_df[-data_split, 3]

target_category1
test_category1

tri3<- knn(train2, test2,cl=target_category1, k=3)
tri3

tri5<- knn(train2, test2,cl=target_category1, k=5)
tri5

tri10<- knn(train2, test2,cl=target_category1, k=10)
tri10

tri15<- knn(train2, test2,cl=target_category1, k=15)
tri15

tri20<- knn(train2, test2,cl=target_category1, k=20)
tri20

tri25<- knn(train2, test2,cl=target_category1, k=25)
tri25
```

 

