---
title: "Exercise16SutowBrett"
author: "Brett Sutow"
date: "10/30/2020"
output: word_document
---

```{r setup, echo=FALSE}
library(FNN)
library(ggplot2)
setwd("/Users/Brett/Desktop/dsc520-master")
datasetcluster_df<- read.csv("/Users/Brett/Desktop/dsc520-master/data/clustering-data.csv")
head(datasetcluster_df)
```

## Answer 1 
a. Below are the scatter plots 


```{r, echo=FALSE}
plot(datasetcluster_df)
ggplot(datasetcluster_df, aes(x= x, y= y)) + geom_point( )+
labs(title= "X vs Y", x="X", y="Y")
```

## Answer 2

b. Below are the k-means results as well as the scatter plots created:
```{r , echo=FALSE}
library('VIM')
library('ggpubr')
library('factoextra')
aggr(datasetcluster_df)
cluster2 <- kmeans(datasetcluster_df[, 1:2], 2)
cluster2
cluster3 <- kmeans(datasetcluster_df[, 1:2],3 )
cluster3
cluster4 <- kmeans(datasetcluster_df[, 1:2],4 )
cluster4
cluster5 <- kmeans(datasetcluster_df[, 1:2],5 )
cluster5
cluster6 <- kmeans(datasetcluster_df[, 1:2],6 )
cluster6
cluster7 <- kmeans(datasetcluster_df[, 1:2],7 )
cluster7
cluster8 <- kmeans(datasetcluster_df[, 1:2],8 )
cluster8
cluster9 <- kmeans(datasetcluster_df[, 1:2],9 )
cluster9
cluster10 <- kmeans(datasetcluster_df[, 1:2],10)
cluster10
cluster11 <- kmeans(datasetcluster_df[, 1:2],11 )
cluster11
cluster12 <- kmeans(datasetcluster_df[, 1:2],12 )
cluster12



fviz_cluster(cluster2, data= datasetcluster_df[, -100],
             palette = c("#2E9FDF", "#00AFBB", "#E7B800"), 
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster3, data= datasetcluster_df[, -100],
             palette = c("#2E9FDF", "#00AFBB", "#E7B800"), 
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster4, data= datasetcluster_df[, -100],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster5, data= datasetcluster_df[, -100],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster6, data= datasetcluster_df[, -100],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster7, data= datasetcluster_df[, -100],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster8, data= datasetcluster_df[, -100],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster9, data= datasetcluster_df[, -100],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster10, data= datasetcluster_df[, -100],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster11, data= datasetcluster_df[, -100],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())

fviz_cluster(cluster12, data= datasetcluster_df[, -100],
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw())
```

## Answer 3 
c. Below are the results for answer 3. These results include the distances requested as well as the average of these distances. There is also a line plot created from this information that helped determine the elbow point. Looking at the graph creating the elbow point would be where k=3 and the average distance is around 88.69. This is where the graph changes shape and has an inflection point. 
```{r , echo=FALSE}
library(philentropy)
distance2<-distance(cluster2$centers, method = "euclidean")
distance2
distance3<-distance(cluster3$centers, method = "euclidean")
distance3
distance4<-distance(cluster4$centers, method = "euclidean")
distance4
distance5<-distance(cluster5$centers, method = "euclidean")
distance5
distance6<-distance(cluster6$centers, method = "euclidean")
distance6
distance7<-distance(cluster7$centers, method = "euclidean")
distance7
distance8<-distance(cluster8$centers, method = "euclidean")
distance8
distance9<-distance(cluster9$centers, method = "euclidean")
distance9
distance10<-distance(cluster10$centers, method = "euclidean")
distance10
distance11<- distance(cluster11$centers, method = "euclidean")
distance11
distance12<- distance(cluster12$centers, method = "euclidean")
distance12

mean2<- mean(distance2)
mean3<-mean(distance3)
mean4<-mean(distance4)
mean5<-mean(distance5)
mean6<-mean(distance6)
mean7<-mean(distance7)
mean8<-mean(distance8)
mean9<-mean(distance9)
mean10<-mean(distance10)
mean11<-mean(distance11)
mean12<-mean(distance12)
mean2 
mean3
mean4
mean5
mean6
mean7
mean8
mean9
mean10
mean11
mean12

averagedf<- data.frame('k' = 2:12, 'means'= c(mean2, mean3, mean4, mean5,mean6, mean7, mean8, mean9, mean10, mean11, mean12))

ggplot(averagedf,aes(x=k, y=means)) +geom_line()

averagedf
```

