---
title: "Exercise_14_SimmersAshley"
author: "Ashley Simmers"
date: "10/25/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logistic Regression


```{r, echo=FALSE}

setwd("C:/Users/ashle/OneDrive/Desktop/DSC_520/DSC_520/dsc520/data")
binary_df <- read.csv("binary-classifier-data.csv")
library(caTools)

split <- sample.split(binary_df, SplitRatio = 0.8)
split
train <- subset(binary_df, split == "TRUE")
test <- subset(binary_df, split == "FALSE")

binary_df$label <- as.factor(binary_df$label)

glm_binary <- glm(label ~ x + y, data = train, family = "binomial")
summary(glm_binary)

res <- predict(glm_binary, test, type = "response")
res <- predict(glm_binary, train, type = "response")

confmatrix <- table(Actual_Value=train$label, Predicted_Value = res > 0.5)
confmatrix

(confmatrix[1,1] + confmatrix[2,2]) / sum(confmatrix)
```
This assignment brought up a lot of questions. What exactly are we trying to predict with our logistic regression? I would think that we would we looking for our dependant variable which I understood would normally be Y. In this I placed label since it is a 0 or 1 binary yes or no. The logistic regression models accuracy was very poor being 58.54%. 


## Nearest Neighbor 


```{r pressure, echo=FALSE}
library(ggplot2)

ggplot(data = binary_df, aes(x = x, y = y, color = label)) + geom_point()
```

Above we can see that there is natural clustering in this data set. 

```{r, echo=FALSE}
library(class)
error.rate <- NULL
predicted.values <- NULL

k.values <- 1:30

for (i in k.values) {
    predicted.values <- knn(train[2:3], test[2:3], train$label, k = i)
    error.rate[i] <- mean(test$label != predicted.values)
}
error.df <- data.frame(error.rate, k.values)

ggplot(data= error.df, aes(x = k.values, y = error.rate)) +
    geom_point() + geom_line(color = "purple") + xlab("K Values") + ylab("Error")


```

Already had created training and test groups. Used them to help determine the right value for K based on rate of error. Lowest error rate is K = 14 and it is 1.4%. 


## Accuracy of Logistic Regression Different

KNN plots data points from your training data and uses test data points to find the nearest neighbors. Logistic regression uses a sigmoid function to determine the probability of a bionary question. 