---
title: "BinaryLogisticRegression"
author: "SumbarajuAditya"
date: "2/7/2021"
output: word_document
bibliography: C:/BU/DSC520/assignment_repo/dsc520/bibliography.bib
---

**Binary logistic regression**
Logistic regression is a classification algorithm and it it also called as logic regression. Binary logistic regression models the relationship between a set of predictors and a binary response variable. (Frost, 2021)
A binary response has only two possible values, such as Yes (1) or No (0). A binary regression model helps us to understand how changes in the predictor values are associated with changes in the probability of an event occurring. [@newtest]

**regression equation**

A regression equation is a statistical model that determined the specific relationship between the predictor variable and the outcome variable. A model regression equation allows you to predict the outcome with a relatively small amount of error.
The mathematical formula of the linear regression can be written as 


$$Y_i=b_0+b_1X_i+ \epsilon_i$$

b0 and b1 are known as the regression beta coefficients or parameters:
b0 is the intercept of the regression line; that is the predicted value when x = 0.
b1 is the slope of the regression line.
e is the error term (also known as the residual errors), the part of y that can be explained by the regression model.

**Replicate the logistic regression example from this link**

https://stats.idre.ucla.edu/r/dae/logit-regression/ 
(Data: https://stats.idre.ucla.edu/stat/data/binary.csv)

```{r}
binary_df <- read.csv("C:/BU/DSC520/assignment_repo/dsc520/data/binary.csv")
head(binary_df)

```
**Dataset Observations**
1. admit - The response variable, admit/don’t admit, it is a binary variable dependent of predictor variable gre,gpa,rank
2. gre- predictor variable
3. gpa - predictor variable
4. rank - predictor variable
5. Rank with 1 have highest prestige and 4 stands lowest and is calculated based on the gpa scores

```{r}
# get the descriptive of dataset
summary (binary_df)
# Standard deviations
sapply (binary_df, sd)
## two-way contingency table of categorical outcome and predictors
xtabs(~admit + rank, data = binary_df)
```
**Logit Model**
1. To estimate the logistic regression model we would be using glm (generalized linear model) function.
2. indicating rank as categorical value
```{r}
binary_df$rank <- factor(binary_df$rank)
logit_glm <- glm(admit ~ gre + gpa + rank, data = binary_df, family = "binomial")
summary(logit_glm)
```

From the coefficients it is evident that 
1. A unit increase in gre increases the log odds of admissions by 0.002264
2. A unit increase in gpa increases the log odds of being admitted by 0.804038

**confidence intervals**

```{r}
confint(logit_glm)
```
test for an overall effect of rank using the wald.test function
```{r}
library(aod)
wald.test(b = coef(logit_glm), Sigma = vcov(logit_glm), Terms = 4:6)
```
The chi-squared test statistic of 20.9, with three degrees of freedom is associated with a p-value of 0.00011 indicating that the overall effect of rank is statistically significant.[@newtest]

**Predicted probabilities**
Predicted probabilities can be computed for both categorical and continuous predictor variables.
To start with Predicted probabilities we need to create a new dataframe with the values we want the independent variables to take on to create our predictions.
```{r}
binary_df1 <- with(binary_df, data.frame(gre = mean(gre), gpa = mean(gpa), rank = factor(1:4)))
binary_df1
```

Let's create a new variable to predict the rank variable using predicted probability (type="response")
```{r}
binary_df1$rankP <- predict(logit_glm, newdata = binary_df1, type = "response")
binary_df1
```
From the above output we see that 
1. The predicted probability of being accepted into a graduate program is 0.5166016 for students from the highest prestige (rank=1) undergraduate institutions
2. 0.1846684 for students from the lowest ranked institutions (rank=4), holding gre and gpa at their means.

**predicted probabilities varying the value of gre and rank**
Let's try to plot 100 values of gre between 200 and 800, at each value of rank (rank=1, 2, 3, and 4).

```{r}
binary_df2 <- with(binary_df, data.frame(gre = rep(seq(from = 200, to = 800, length.out = 100),
    4), gpa = mean(gpa), rank = factor(rep(1:4, each = 100))))

binary_df3 <- cbind(binary_df2, predict(logit_glm, newdata = binary_df2, type = "link",
    se = TRUE))
binary_df3 <- within(binary_df3, {
    PredictedProb <- plogis(fit)
    LL <- plogis(fit - (1.96 * se.fit))
    UL <- plogis(fit + (1.96 * se.fit))
})

## view first few rows of final dataset
head(binary_df3)
```

**ggplot2**

```{r}
library(ggplot2)
ggplot(binary_df3, aes(x = gre, y = PredictedProb)) + geom_ribbon(aes(ymin = LL,
    ymax = UL, fill = rank), alpha = 0.2) + geom_line(aes(colour = rank),
    size = 1)
```

Is our model with predictors fits significantly better than a model with just an intercept? Lets test it.

```{r}
with(logit_glm, null.deviance - deviance)
with(logit_glm, df.null - df.residual)
```

#Claculate P-Value
```{r}
options(scipen=10000)
with(logit_glm, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE))
```
The chi-square of 41.45903 with 5 degrees of freedom and an associated p-value of less than 0.001 tells us that our model as a whole fits significantly better than an null model.
[@newtest]

**references:**
* Introduction to SAS. UCLA: Statistical Consulting Group [@newtest]
* Frost, J. (2021). Binary logistic regression. Retrieved February 08, 2021, from https://statisticsbyjim.com/glossary/binary-logistic-regression/
