---
title: "Exercise 6 - Regression with R"
author: "Amy Femal"
date: "7/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/15626/Desktop/dsc520/data')
```

Load the `data/r4ds/heights.csv` to
```{r}
heights_df <- read.csv("r4ds/heights.csv")
```
Load the ggplot2 library
```{r}
library(ggplot2)
```
Fit a linear model using the `age` variable as the predictor and `earn` as the outcome
```{r}
age_lm <-  lm(heights_df$earn~heights_df$age)
```

View the summary of your model using `summary()`
```{r}
summary(age_lm)
```

Creating predictions using `predict()`
```{r}
age_predict_df <- data.frame(earn = predict(age_lm, heights_df), age=heights_df$age)
```

Plot the predictions against the original data
```{r}
ggplot(data = heights_df, aes(y = earn, x = age)) +
  geom_point(color='blue') +
  geom_line(color='red',data = age_predict_df, aes(y= earn, x=age))
```
```{r}
mean_earn <- mean(heights_df$earn)
```
Corrected Sum of Squares Total
```{r}
sst <- sum((mean_earn - heights_df$earn)^2)
```
Corrected Sum of Squares for Model
```{r}
ssm <- sum((mean_earn - age_predict_df$earn)^2)
```
Residuals
```{r}
residuals <- heights_df$earn - age_predict_df$earn
```
Sum of Squares for Error
```{r}
sse <- sum(residuals^2)
```
R Squared R^2 = SSM\SST
```{r}
r_squared <- ssm/sst
```
Number of observations
```{r}
n <- length(heights_df$age)
```
Number of regression parameters
```{r}
p <- 2
```
Corrected Degrees of Freedom for Model (p-1)
```{r}
dfm <- p - 1
```
Degrees of Freedom for Error (n-p)
```{r}
dfe <- n - p
```
Corrected Degrees of Freedom Total:   DFT = n - 1
```{r}
dft <- n - 1
```
Mean of Squares for Model:   MSM = SSM / DFM
```{r}
msm <- ssm/dfm
```
Mean of Squares for Error:   MSE = SSE / DFE
```{r}
mse <- sse/dfe
```
Mean of Squares Total:   MST = SST / DFT
```{r}
mst <- sst/dft
```
F Statistic F = MSM/MSE
```{r}
f_score <- msm/mse
```

Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r}
adjusted_r_squared <- 1 - (1 - r_squared)*(n - 1)/(n - p)
```
Calculate the p-value from the F distribution
```{r}
p_value <- pf(f_score, dfm, dft, lower.tail=F)
```
We can compare the $R^2$, Adjusted $R^2$ and F statistics.
```{r}
names(summary(age_lm))
```

```{r}
r_squared
summary(age_lm)$r.squared
```
```{r}
adjusted_r_squared
summary(age_lm)$adj.r.squared
```

```{r}
f_score
summary(age_lm)$fstatistic
```
We can see we get the same results.
